Self-Supervised Transfer Learning From 2D Images to 3D Point Clouds

### 2022-04-01 doc/paper.tex
Adds pretrained and supervised results
adds first part of methods section

%%%%%%%%% TITLE - PLEASE UPDATE
\title{Exploring Self-Supervised Learning for Indoor 3D scenes}

@@

\section{Method}
\label{sec:method}

\section{Feature Extractors}
\label{sec:method:featuresExtractors}

\section{Extracting 2D features}
\label{sec:method:features2d}

\section{Transfering 2D features}
\label{sec:method:features2D3D}

@@

\begin{table*}[t!]
    \centering
    \resizebox{\textwidth}{!}{
        \begin{tabular}{ c | c c c c c c c c c c c c c | c }
                                        & ceiling        & floor          & wall           & beam & column & window & door           & table          & chair          & sofa           & bookcase       & board & clutter        & mIOU           \\
            \hline
            Scratch                     & 91.53          & 96.79          & 81.25          & 0.0  & 29.81  & 56.92  & 64.77          & 74.43          & 86.35          & 55.24          & 68.38          & 79.7  & 55.65          & 64.68          \\
            Supervised                  &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
            \hline
            PointContrast               &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
            PointContrast - No Overhang &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
            CME                         & \textbf{91.07} & \textbf{96.74} & \textbf{82.84} & 0    & 28.38  & 55.98  & \textbf{70.34} & \textbf{75.01} & \textbf{87.34} & \textbf{63.24} & \textbf{71.25} & 75.01 & \textbf{56.42} & \textbf{65.67} \\
        \end{tabular}
    }
    \caption{S3DIS results using ContrastiveSceneContexts Code}
    \label{table:s3disResults}
\end{table*}

@@

>>>>

%%%%%%%%% TITLE - PLEASE UPDATE
\title{Self-Supervised Transfer Learning From 2D Images to 3D Point Clouds}

@@

\section{Self-Supervised Contrastive learning}
\label{sec:contrastiveLearning}

The goal of self-supervised transfer learning is to learn from unlabeled data using a self-supervised objective and use the initialized backbone parameters to improve performance of downstream tasks. Specifically, contrastive learning uses an objective function that encourages the outputs of a model on similar datapoints to be close together while keeping those of dissimilar datapoints to be far apart.

Defining what datapoints are similar and dissimilar depends on the pretext task. Image colorization for example defines color pixels and their grayscale equivalent as similar and all other pixels as dissimilar. In this work, we adopt the instance discrimination task as in \cite{} whereby each input individually is taken to belong to its own class.

Given a sampled point $\mathbf{x}$, positive points are sampled according to a positive distribution $\mathbf{x}^{+} \sim p^{+}( \cdot | \mathbf{x})$ which is commonly defined as a set of transformations $t$ of the original point. Often, positive pairs are derived by applying two different augmentations to the input point, $x = t(x), \quad x^{+} = t^{'}(x)$. Conversely, negative points are sampled according to $\mathbf{x}^{-} \sim p^{-}( \cdot | \mathbf{x})$ which in the context of an instance discriminative task means that they are taken as any other point not derived from the input point, regardless of augmentation.

As for the final loss, we adopt InfoNCE (Info Noise Contrastive Estimation) loss which is a negative log probability of a point's similarity to its positive sample conditioned by the similarity of that point to all the negative samples. It is defined as:
\begin{equation}
    \mathcal{L}_{i} = -\log \frac{\exp(x_{i} \cdot x^{+}_{i} / \tau)}{\exp(x_{i} \cdot x^{+}_{i} / \tau) + \sum^{K}_{j}\exp(x_{i} \cdot x^{-}_{j} / \tau)}
    \label{eq:contrastive_loss}
\end{equation}

Where $\tau \in (0,1]$ and represents the temperature parameter that controls the smoothness of the latent representations. Similarity between features is computed as the dot product.

\subsection{Feature Extraction}
\label{sec:method:featuresExtractors}

Then we need an encoder and decoder.

Throw away the end of the

Use a ResNet and U-net following the same structure as was defined in...

\subsection{Extracting 2D features}
\label{sec:method:features2d}

How the loss was formulated in 2D.

What augmentations, cropping the image in two places, only using the points on the overlap regions.

Add figure of what your feature extractor looks like

\subsection{Transfering 2D features}
\label{sec:method:features2D3D}

How the loss was formulated in the 2D to 3D transfer.

@@

\begin{table*}[t!]
    \centering
    \resizebox{\textwidth}{!}{
        \begin{tabular}{ c | c c c c c c c c c c c c c | c }
                                        & ceiling         & floor          & wall           & beam          & column         & window         & door           & table          & chair           & sofa           & bookcase       & board          & clutter        & mIOU                    \\
            \hline
            Scratch                     & 91.53           & \textbf{96.79} & 81.25          & 0             & 29.81          & 56.92          & 64.77          & 74.43          & 86.35           & 55.24          & 68.38          & 79.7           & 55.65          & 64.68                   \\
            Supervised                  & \textbf{91.81}  & 96.74          & \textbf{84.11} & \textbf{0.21} & \textbf{34.68} & \textbf{57.64} & \textbf{81.53} & \textbf{77.63} & \textbf{89.82}  & \textbf{85.76} & \textbf{76.03} & \textbf{79.27} & \textbf{57.67} & \textbf{70.22} (+5.54)  \\
            \hline
            PointContrast               & \textbf{92.118} & \textbf{96.93} & 82.2           & 0             & 24.94          & 53.78          & \textbf{76.91} & \textbf{75.94} & \textbf{87.929} & \textbf{70.27} & 71.196         & \textbf{77.76} & 56.36          & \textbf{66.641} (+1.73) \\
            PointContrast - No Overhang &                 &                &                &               &                &                &                &                &                 &                &                &                &                &                         \\
            CME                         & 91.07           & 96.74          & \textbf{82.84} & 0             & \textbf{28.38} & \textbf{55.98} & 70.34          & 75.01          & 87.34           & 63.24          & \textbf{71.25} & 75.01          & \textbf{56.42} & 65.67 (+0.99)           \\
        \end{tabular}
    }
    \caption{S3DIS results using ContrastiveSceneContexts Code}
    \label{table:s3disResults}
\end{table*}

@@

Things to include in results
\begin{itemize}
    \item S3DIS semantic segmentation
    \item S3DIS instance segmentation
    \item Scannet semantic segmentation
    \item Visualization of features in 2d
    \item Visualization of features in 3D
    \item Training and validation curves
    \item Ablation studies
\end{itemize}
